Commutative Character Sheaves

### 2015-11-23 followup.tex
Catching up...
Work on the section reviewing types. Models is under construction.

@@

%\subsection{Geometric types for supercuspidal representations}

\newcommand{\red}{^{\operatorname{red}}}

\subsection{Review of Yu's types}

A Yu type is a pair $(\,^\circ K,\,^\circ \rho)$, where $\,^\circ K$ is a compact group and $\,^\circ \rho$ is a representation of $\,^\circ K$, with certain properties. 
One of the main results of \cite{yu:models} is that there is a canonical affine smooth group scheme $\underline{G}$ of finite type over $\OK$ such that $\,^\circ K = \underline{G}(\OK)$ \cite{yu:models}*{Prop 10.2}.
As explained in \cite{yu:models}*{\S 10.4}, the
the group scheme $\underline{G}$ comes with the following auxillary data:
\begin{enumerate}
\item[(Ji)]
	affine smooth group schemes $\underline{J}^i$, for $i=0, 1, \ldots, d$, together with morphisms $\underline{J}^i \to \underline{G}$ such that \[\prod_{i=0}^d \underline{J}^i(\OK) = \underline{G}(\OK),\] where we identify $\underline{J}^i(\OK)$ with its image under $\underline{J}^i \to \underline{G}$, and such that $\prod_{i=j+1}^{d} \underline{J}^i(\OK)$ is normal in $\underline{G}(\OK)$, for $j=1, \ldots, d$;
\item[(Gj)] affine smooth group schemes $\underline{G}^{j}$, for $j=1, \ldots, d$, together with morphisms $\underline{G}^{j}\to \underline{G}$ such that 
\[
\underline{G}^{j}(\OK) = \prod_{i=0}^{j} \underline{J}^i(\OK),
\] 
where we identify $\underline{J}^i(\OK)$ with its image under $\underline{J}^i \to \underline{G}$ and $\underline{G}^{j}(\OK)$ with its image under $\underline{G}^j \to \underline{G}$.
\end{enumerate}

The representation $\,^\circ \rho$ of $\underline{G}(\OK)$ is build from the following data:
\begin{enumerate}
\item
	a representation $\sigma$ of $\underline{G}_s\red(\Fq)$;
\item
if $d \geq 2$, for $j=1,\ldots, d-1$, 
\begin{enumerate}
\item
a representation $\beta_j$ of $H^j(\Fq)$, where $H^j$ is Heisenberg group arising as a quotient $(\underline{J}^{j})_s \to H^j$,
\item
a representation $\kappa_j$ of $(\underline{G}^{j-1})_s\red(\Fq)$
\item
a linear quasicharacter $\phi_{j-1}$ of $\underline{G}^{j-1}(\OK)$;
\end{enumerate}
\item
	a linear quasicharacter $\phi$ of $\underline{G}(\OK)$.
\end{enumerate}
From this data, the representation $\,^\circ \rho$ is built as follows, as explained in \cite{yu:models}*{\S 10.5}.
\begin{enumerate}
\item 
	Let $\dot{\sigma}$ be the representation of $\underline{G}_s(\Fq)$ obtained by pulling back $\sigma$ along $\underline{G}_s(\Fq) \to \underline{G}_s\red(\Fq)$. 	
	Let $\ddot{\sigma}$ be the representation of $\underline{G}(\OK)$ obtained by pullback along $\underline{G}(\OK) \to \underline{G}_s(\Fq)$.
\item For $j=1,\ldots, d-1$, 
\begin{enumerate}
\item
	let $\dot{\beta}_j$ be the representation of $\underline{J}^{j}_s(\Fq)$ obtained by pulling back $\beta_j$ along $\underline{J}^{j}_s(\Fq) \to H^j(\Fq)$ and let $\ddot{\beta}_j$ be . . . 
;
\item 
	let $\dot{\kappa}_j$ be the representation of $\underline{G}^{j-1}_s(\Fq)$ obtained by pulling back $\kappa_j$ along $\underline{G}^{j-1}_s(\Fq) \to (\underline{G}^{j-1})_s\red(\Fq)$ and let $\ddot{\kappa}_j$ . . ;
\item
	let $\phi_{j-1}$ be the linear quasicharacter of $\underline{G}^{j-1}(\OK)$ as above. 
\end{enumerate}
\item Let $\phi$ be the linear quasicharacter of $\underline{G}(\OK)$, as above.
\end{enumerate}
Using the decomposition $\underline{G}^j(\OK) = \underline{G}^{j-1}(\OK) \times \underline{J}^j(\OK)$, combine $\ddot{\beta}_j$, $\ddot{\kappa}_j$  and $\phi_j$ to form the representation $(\phi_{j-1}\otimes \ddot{\kappa}_j)\times \ddot{\beta}_j$ of $\underline{G}^{j}(\OK)$, for $j=1,\ldots, d-1$. 
	Since $\prod_{i=j+1}^{d} \underline{J}^i(\OK)$ is normal in $\underline{G}(\OK)$, this determines a representation $\delta_j$ of $\underline{G}(\OK)$, for $j=1,\ldots, d-1$.
Then 
\[
\,^\circ \rho = \ddot{\sigma} \otimes \delta_1 \cdots \otimes \delta_{d-1} \otimes \phi.
\]

\subsection{Geometrization of types}

Let $(\underline{G}(\OK),\,^\circ \rho)$ be a Yu type.
Set $G = \Gr(\underline{G})$, the Greenberg transform of $\underline{G}$.
Recall that $G$ is a group scheme over $\Fq$ and comes equipped with

\begin{proposition}\label{prop:types}
Let $(\underline{G}(\OK),\rho)$ be a Yu type.
With notation as above, suppose $\pi_0(\underline{G}_s\red)$ is cyclic and that $\pi_0((\underline{G}^{j-1})_s\red)$ is cyclic for $j=1,\ldots, d-1$.
Then there is a virtual admissible $\ell$-adic sheaf $\mathcal{F}$ on $G$ such that $\trFrob{\mathcal{F}} = \trace \rho$ under the canonical isomorphism $G(\Fq) = \underline{G}(\OK)$.
\end{proposition}

>>>>

\newcommand{\red}{^{\operatorname{red}}}
\newcommand{\Sp}{{\operatorname{Sp}}}

A Yu type is a pair $(\,^\circ K,\,^\circ \rho)$, where $\,^\circ K$ is a compact group and $\,^\circ \rho$ is a representation of $\,^\circ K$, with certain properties. 
In Section~\ref{ssec:review} we recall properties of $(\,^\circ K,\,^\circ \rho)$ which we allow us, in Section~\ref{ssec:geometrization}, to find a pair $(G, \mathcal{F})$ where $G$ is a smooth group scheme over $\Fq$ and $\mathcal{F}$ is a virtual constructible sheaf $\ell$-adic sheaf on $G$ from which the character of $\,^\circ \rho$ can be recovered using the trace of Frobenius.

\subsection{Review of Yu's types}\label{ssec:review}

Every Yu type $(\,^\circ K,\,^\circ \rho)$ may be constructed directly from the following information, derived from a Yu datum:
\begin{enumerate}
\item compact open subgroups $J^i\subset \,^\circ K$, for $i=0, \ldots d$, such that \[\,^\circ K = J^0\cdots J^d;\]
\item a continuous representation $\sigma_0$ of $J^0$;
\item a quasicharacter $\varphi_i : J^i \to \CC^\times$, for $i=0, \ldots d$.
\end{enumerate}
These representations $(\sigma_0, \varphi_0, \ldots , \varphi_d)$ enjoy certain properties, which allow Yu to construct, for $i=0, \ldots, d-1$, further:
\begin{enumerate}
\item a quotient $J^{i+1} \to V_{i+1}$ to a finite abelian group, and the structure of a $\Fq$-vector space on $V_{i+1}$, compatible with the abelian group structure above;
\item a symplectic pairing $\langle\ ,\ \rangle_{i+1} : V_{i+1}\times V_{i+1} \to C_{i+1}$, where $C_{i+1}$ is a one-dimensional vector space over $\Fq$;
\item an additive character $\psi_{i+1} : C_{i+1} \to \EEx$;
\item a group homomorphism $\,^\circ f_i : J^0 \cdots J^i \to \Sp(V_{i+1})$;
\item a quotient $J^{i+1} \to H^{i+1}$, where $H^{i+1}$ is a Heisenberg $p$-group, together with an isomorphism $j_{i+1} : H^{i+1} \to V_{i+1}^\sharp$, where $V_{i+1}^\sharp$ is the Heisenberg group determined by $(V_{i+1}, \langle\ ,\ \rangle_{i+1}, C_{i+1}, \psi_{i+1})$.
\end{enumerate}
The pair $(f_i, j_{i+1})$ is a symplectic action of $J^0\cdots J^i$ on $H^{i+1}$, so it defines a group homomorphism 
\begin{equation}\label{eq:symaction}
\begin{tikzcd}
h_{i+1} : (J^0\cdots J^i)\ltimes J^{i+1} \arrow{r} & (J^0\cdots J^i)\ltimes H^{i+1} \arrow{r} & \Sp(V_{i+1})\ltimes V_{i+1}^\sharp.
\end{tikzcd}
\end{equation}
%

From this information, Yu constructs a representation $\,^\circ \rho_i$ of $J^0\cdots J^i$, for each $i=0, \ldots, d$, such that $\,^\circ \rho_d = \,^\circ \rho$. We will need some detail about this construction, so we will review it here.
%Let $\varpi_1$ be the Heisenberg representation of the Heisenberg group $V_1^\sharp$, with central character $\psi_1$.
%Let $\varrho_1$ be the Weil representation of $\Sp(V_1)$ with the same representation space as $\omega_1$.
Let $\omega_{i+1}$ be the Heisenberg-Weil representation of the Jacobi group $\Sp(V_{i+1})\ltimes V_{i+1}^\sharp$, with the same representation space as the Heisenberg representation of the Heisenberg group $V_{i+1}^\sharp$.
\[
\begin{tikzcd}
\  & J^{i+1} \arrow{d} & \\
Sp(V_{i+1}) \ltimes V_{i+1}^\sharp & \arrow{l}[swap]{h_1} (J^0 \cdots J^i) \ltimes J^{i+1} \arrow{r} \arrow{d} & J^0 \cdots J^i J^{i+1}\\
& (J^0 \cdots J^i) & 
\end{tikzcd}
\]
Then $h_1^*(\omega_1)$ is a representation of $J^0\ltimes J^1$, determined by the quasicharacter $\varphi_1$ of $J^1$. 
Write $\inf(\sigma_0\otimes \varphi_0)$ for the representation of $J^0 \ltimes J^1$ obtained by pulling back $\sigma_0\otimes \varphi_0$ along $J^0 \ltimes J^1 \to J^0$. 
By \cite{yu:construction}*{}, the representation 
$h_1^*(\omega_1) \otimes \inf(\sigma_0\otimes \varphi_0)$ of $J^0 \ltimes J^1$ is trivial on $(J^0 \ltimes 1) \cap (1 \ltimes J^1)$, so $h_1^*(\omega_1) \otimes \inf(\sigma_0\otimes \varphi_0)$ descends to a representation $\,^\circ \rho_1$ of $J^0 J^1$.  
%
As above, the representation 
$h_{2}^*(\omega_{2}) \otimes \inf(\,^\circ \rho_1)$ of $J^0 \ltimes J^1$ is trivial on $(J^0 J^1 \ltimes 1) \cap (1 \ltimes J^2)$, so $h_2^*(\omega_2) \otimes \inf(\,^\circ \rho_1)$ descends to a representation $\,^\circ \rho_2$ of $J^0 J^1 J^2$.  
Continuing in this way, one constructs the representation $\,^\circ \rho_d$ of $J^0\cdots J^d$.

The character of $\,^\circ \rho_d$ on $J^0\cdots J^d$ satisfies
\[
\trace(\,^\circ \rho_d)(k_0 \cdots k_d) = \sigma_0(k_0) \left( \varphi_0(k_0) \cdots \varphi_d(k_d) \right) ,
\]
with ...

\subsection{Smooth models}\label{ssec:models}

The first step in the geometrization of the type $(\,^\circ K,\,^\circ \rho)$ is to replace $\,^\circ K$ with a group scheme $G$ over $\Fq$ so that $G(\Fq) = \,^\circ K$, canonically. 
By \cite{yu:models}*{Prop 10.2} there is a canonical affine smooth group scheme $\underline{G}$ of finite type over $\OK$ such that \[\underline{G}(\OK)= \,^\circ K\] 
For every positive integer $n$, let $G_n$ be the Greenberg transform of $\underline{G} \times_{\Spec{\OK}} \Spec{\OK/\mathfrak{m}_K^n}$, where $\mathfrak{m}_K$ is the maximal ideal in $\OK$.
Then $G_n$ is a smooth group scheme over $\Fq$ and $G_n(\Fq) = \underline{G}(\OK/\mathfrak{m}_K^n)$, canonically.
Let $G$ be the Greenberg transform of $\underline{G}$.
Then $G$ is a group scheme over $\Fq$ and $G(\Fq) = \underline{G}(\OK) = \,^\circ K$, canonically.

In order to complete the geometrization of the type $(\,^\circ K,\,^\circ \rho)$, we will need more smooth models.
Fortunately, as above, Yu has already done the heavy work.
%
As explained in \cite{yu:models}*{\S 10.4}, the
the group scheme $\underline{G}$ comes with the following auxillary data:
\begin{enumerate}
\item[(Ji)]
	affine smooth group schemes $\underline{J}^i$, for $i=0, 1, \ldots, d$, together with morphisms $\underline{J}^i \to \underline{G}$ such that \[\underline{J}^i(\OK) = J^i,\]
\item[(Gj)] affine smooth group schemes $\underline{G}^{j}$, for $j=0, \ldots, d$, together with morphisms $\underline{G}^{j}\to \underline{G}$ such that 
\[
\underline{G}^{j}(\OK) = J^0\cdots J^i.
\] 
\end{enumerate}

Moreover, the quotient

\subsection{Geometrization}\label{ssec:geometrization}

\begin{proposition}\label{prop:types}
Let $(\,^\circ K,\,^\circ \rho)$ be a Yu type.
With notation as above, suppose $\pi_0(\underline{G}_s\red)$ is cyclic and that $\pi_0((\underline{G}^{j-1})_s\red)$ is cyclic for $j=1,\ldots, d-1$.
Then there is a virtual admissible $\ell$-adic sheaf $\mathcal{F}$ on $G$ such that $\trFrob{\mathcal{F}} = \trace \rho$ under the canonical isomorphism $G(\Fq) = \underline{G}(\OK)$.
\end{proposition}

@@
